---
title: "'RCTrep: An R package for replicating treatment effect estimates of a randomized"
  control trial using observational data: A case study'
output:
  html_document:
    df_print: paged
---
# Working example 1: Replicating QUASAR trial using NCR data
# Import data
```{r}
library(RCTrep)
library(dplyr)
source.data <- RCTrep::ncr
target.data <- RCTrep::quasar.agg
head(target.data)
head(source.data)
```

# Initialze source.obj for NCR data
## Arguments 
```{r}
TEstimator <- "G_computation"
outcome_formula <- vitstat ~ combined_chemo + Stage2 + age + pT + BRAF + male + 
                   combined_chemo:Stage2 +combined_chemo:age + combined_chemo:pT + 
                   combined_chemo:male + combined_chemo:pT:Stage2 + combined_chemo:BRAF
vars_name <- list(confounders_treatment_name=c("Stage2","age","pT","BRAF","male"),
                  treatment_name=c('combined_chemo'),
                  outcome_name=c('vitstat')
                  )
```

## Initialize 
```{r, warning=FALSE}
#source.data$combined_chemo <- as.numeric(as.character(source.data$combined_chemo))
source.obj <- TEstimator_wrapper(
  Estimator = TEstimator,
  data = source.data,
  name = "ncr",
  vars_name = vars_name,
  outcome_method = "glm",
  outcome_formula = outcome_formula,
  data.public = TRUE
)
```

```{r}
source.obj$data
```


## summarize data distribution
```{r}
source.obj$diagnosis_t_overlap(stratification = c("Stage2","pT"),stratification_joint = TRUE)
```

```{r}
source.obj$diagnosis_y_overlap(stratification = c("Stage2","age","pT"),stratification_joint = TRUE)
```

## diagnosis estimates
```{r}
source.obj$plot_CATE(stratification = c("Stage2","pT"))
```

## diagnosis model fit
```{r}
source.obj$summary()
```

# Inialize target.obj for RCT data (QUASAR trial)
```{r}
target.obj <- Synthetic_TEstimator$new(dim(source.data)[1], target.data, name="QUASAR trial")
```

# Inialize target.obj.rep
```{r}
confounders_sampling <- c("Stage2","male","age")
source.obj.rep <- SEstimator_wrapper(estimator="exact_pp",
                                     target.obj=target.obj,
                                     source.obj=source.obj,
                                     confounders_sampling=confounders_sampling)
```
## Implement RCT replication
```{r,warning=FALSE}
source.obj.rep$EstimateRep(stratification = c("Stage2","male","age"), stratification_joint = FALSE)
```

## Plot
```{r}
fusion <- Summary$new(target.obj,
                      source.obj,
                      source.obj.rep)
fusion$plot()
```


# The Netherlands vs. Korea
```{r}
target.data <- RCTrep::korea
```

## Initialize target.obj for korea
```{r,warning=FALSE}
target.obj <- TEstimator_wrapper(
  Estimator = TEstimator,
  data = target.data,
  name = "korea",
  vars_name = vars_name,
  outcome_method = "glm",
  outcome_formula = outcome_formula,
  data.public = FALSE
)
```

```{r}
confounders_sampling <- c("Stage2","pT","age")
source.obj.rep <- SEestimator_wrapper(estimator="exact_pp",
                                      target.obj=target.obj,
                                      source.obj=source.obj,
                                      confounders_sampling=confounders_sampling)

```

```{r}
source.obj.rep$diagnosis(stratification = c("Stage2","pT","age"))
```


## Implement RCT replication
```{r}
source.obj.rep$EstimateRep(stratification = c("Stage2","age"), stratification_joint = TRUE)
```

## Plot
```{r}
fusion <- Summary$new(target.obj,
                      source.obj,
                      source.obj.rep)
fusion$plot()
```










